Assessing exposure to Atlantic Basin tropical storms in United States counties

### 2016-06-17 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Edit paper draft

---
title: Assessing exposure to Atlantic Basin tropical storms in United States counties
author:
  - name: Joshua Ferreri
    email: joshua.m.ferreri@gmail.com
    affiliation: Colorado State University
  - name: Meilin Yan
    email: alice@example.com
    affiliation: Colorado State University
  - name: Mohammad Z. Al-Hamdan
    email: alice@example.com
    affiliation: NASA Marshall Space Flight Center
  - name: William L. Crosson
    email: alice@example.com
    affiliation: NASA Marshall Space Flight Center
  - name: Seth Guikema
    email: alice@example.com
    affiliation: University of Michigan
  - name: Roger D. Peng
    email: alice@example.com
    affiliation: Johns Hopkins Bloomberg School of Public Health
  - name: G. Brooke Anderson
    email: brooke.anderson@colostate.edu
    affiliation: Colorado State University
    footnote: Corresponding Author
address:
  - code: Colorado State University
    address: Department of Environmental and Radiological Health Sciences, Lake Street, Fort Collins, CO, Zip
  - code: NASA Marshall Space Flight Center
    address: Universities Space Research Association, Street, Huntsville, AL, Zip
  - code: University of Michigan
    address: Department, Street, City, State, Zip
  - code: Johns Hopkins Bloomberg School of Public Health
    address: Department of Biostatistics, Street, Baltimore, MD, Zip
abstract: |
  There are many important applications for having county-level estimates of exposure to tropical storms over many years. For example, ... . Current approaches include ... but have the following limitations ... .

@@

```{r echo = FALSE}
library(hurricaneexposure)
library(knitr)
```

@@

Given its high resolution and consistancy with rain gage values, [@Villarini2011] recommends the use of Stage IV precipitation data for the analysis of rainfall distribution in landfalling tropical cyclones. However, the Stage IV dataset only provides data since 2001 [@Lin2005] and is therefore inadequate for use in this analysis.

@@

```{r echo = FALSE}
library(hurricaneexposure)
map_rain_exposure("Floyd-1999", rain_limit = 75, dist_limit = 500,
                      days_included = -1:1)
```

@@

In comparison to the storm track, the areas of extreme rain and extreme wind can vary. For example, storms undergoing an extratropical transition can bring heavy rains to the left of the storm's center track, while extreme winds are more common to the right of the track [@Halverson2015, @Grabich2016 GBA-- I think it would be better to just use meteorological papers as references for this point. The Halverson paper is great, but the Grabich paper I think might just be a secondary, rather than primary, reference for this point, so I think we could probably find a more direct reference]. Wind speeds are typically more severe to the right of the storm's track because here the counter-clockwise cyclonic [right word?] winds are moving in the same direction as the forward motion of the storm, creating an additive effect for total windspeeds [citation]. In contrast, rains can be heaviest to the left of the storm's track, especially in cases when the storm interacts with a downstream ridge [@Atallah2007].

@@

For example,

```{r echo = FALSE}
map_rain_exposure("Floyd-1999", rain_limit = 75, dist_limit = 500,
                      days_included = -1:1)
```

>>>>

---
title: Assessing exposure to Atlantic Basin tropical storms in United States counties
author:
  - name: Joshua Ferreri
    email: joshua.m.ferreri@gmail.com
    affiliation: Colorado State University
  - name: Meilin Yan
    email: meilin.yan@colostate.edu
    affiliation: Colorado State University
  - name: Mohammad Z. Al-Hamdan
    email: mohammad.alhamdan@nasa.gov
    affiliation: NASA Marshall Space Flight Center
  - name: William L. Crosson
    email: bcrosson@usra.edu
    affiliation: NASA Marshall Space Flight Center
  - name: Seth Guikema
    email: sguikema@umich.edu
    affiliation: University of Michigan
  - name: Roger D. Peng
    email: rdpeng@jhu.edu
    affiliation: Johns Hopkins Bloomberg School of Public Health
  - name: G. Brooke Anderson
    email: brooke.anderson@colostate.edu
    affiliation: Colorado State University
    footnote: Corresponding Author
address:
  - code: Colorado State University
    address: Department of Environmental and Radiological Health Sciences, Lake Street, Fort Collins, CO, Zip
  - code: NASA Marshall Space Flight Center
    address: Universities Space Research Association, 320 Sparkman Dr., Huntsville, AL, 35805
  - code: University of Michigan
    address: Department of Industrial and Operations Engineering, 1205 Beal Ave., Ann Arbor, MI, 48109
  - code: Johns Hopkins Bloomberg School of Public Health
    address: Department of Biostatistics, 615 North Wolfe Street, Baltimore, MD, 21205
abstract: |
  There are many important applications for having county-level estimates of exposure to tropical storms over many years. For example, ... . Current approaches include ... but have the following limitations ... .

@@

```{r echo = FALSE}
library(hurricaneexposure)
library(knitr)
library(gridExtra)
```

@@

Given its better consistancy with rain gage values compared to NLDAS precipitation data [GBA-- you had something in here about resolution, as well. Does it have a better resolution than the NLDAS data? I think we probably want to focus on comparing this source with NLDAS, and we don't need to put much about advantages that it has that are very similar to NLDAS], [@Villarini2011] recommends Stage IV precipitation data for the analysis of rainfall distribution in landfalling tropical cyclones. However, the Stage IV dataset has only been archived since 2002 [@Lin2005] and cannot be used to analyze exposure to tropical storms for studies that include storms before 2002.

@@

```{r echo = FALSE, fig.width = 6, fig.height = 8}
a <- map_counties(storm = "Floyd-1999", metric = "rainfall")
a <- map_tracks(storms = "Floyd-1999", plot_object = a, plot_points = FALSE)
b <- map_counties(storm = "Floyd-1999", metric = "distance")
b <- map_tracks(storms = "Floyd-1999", plot_object = b, plot_points = FALSE)
grid.arrange(a, b, ncol = 1)
```

@@

In comparison to the storm track, the areas of extreme rain and extreme wind can vary. For example, storms undergoing an extratropical transition can bring heavy rains to the left of the storm's center track [@Halverson2015], while extreme winds are more common to the right of the track [@Halverson2015, @Grabich2016 GBA-- I think it would be better to just use meteorological papers as references for this point. The Halverson paper is great, but the Grabich paper I think might just be a secondary, rather than primary, reference for this point, so I think we could probably find a more direct reference]. Wind speeds are typically more severe to the right of the storm's track because here the counter-clockwise cyclonic [right word?] winds are moving in the same direction as the forward motion of the storm, creating an additive effect for total windspeeds [citation]. In contrast, rains can be heaviest to the left of the storm's track, especially in cases when the storm interacts with a downstream ridge [@Atallah2007].

@@

### 2016-06-17 writing/DraftExposurePaper/DraftExposurePaper.tex
Edit paper draft

    \author[Colorado State University]{Meilin Yan}
   \ead{alice@example.com}

    \author[NASA Marshall Space Flight Center]{Mohammad Z. Al-Hamdan}
   \ead{alice@example.com}

    \author[NASA Marshall Space Flight Center]{William L. Crosson}
   \ead{alice@example.com}

    \author[University of Michigan]{Seth Guikema}
   \ead{alice@example.com}

    \author[Johns Hopkins Bloomberg School of Public Health]{Roger D. Peng}
   \ead{alice@example.com}

    \author[Colorado State University]{G. Brooke Anderson\corref{c1}}
   \ead{brooke.anderson@colostate.edu} 
   \cortext[c1]{Corresponding Author}
      \address[Colorado State University]{Department of Environmental and Radiological Health Sciences, Lake
Street, Fort Collins, CO, Zip}
    \address[NASA Marshall Space Flight Center]{Universities Space Research Association, Street, Huntsville, AL, Zip}
    \address[University of Michigan]{Department, Street, City, State, Zip}
    \address[Johns Hopkins Bloomberg School of Public Health]{Department of Biostatistics, Street, Baltimore, MD, Zip}

@@

For example,

\includegraphics{DraftExposurePaper_files/figure-latex/unnamed-chunk-3-1.pdf}

>>>>

    \author[Colorado State University]{Meilin Yan}
   \ead{meilin.yan@colostate.edu}

    \author[NASA Marshall Space Flight Center]{Mohammad Z. Al-Hamdan}
   \ead{mohammad.alhamdan@nasa.gov}

    \author[NASA Marshall Space Flight Center]{William L. Crosson}
   \ead{bcrosson@usra.edu}

    \author[University of Michigan]{Seth Guikema}
   \ead{sguikema@umich.edu}

    \author[Johns Hopkins Bloomberg School of Public Health]{Roger D. Peng}
   \ead{rdpeng@jhu.edu}

    \author[Colorado State University]{G. Brooke Anderson\corref{c1}}
   \ead{brooke.anderson@colostate.edu} 
   \cortext[c1]{Corresponding Author}
      \address[Colorado State University]{Department of Environmental and Radiological Health Sciences, Lake
Street, Fort Collins, CO, Zip}
    \address[NASA Marshall Space Flight Center]{Universities Space Research Association, 320 Sparkman Dr., Huntsville,
AL, 35805}
    \address[University of Michigan]{Department of Industrial and Operations Engineering, 1205 Beal Ave., Ann
Arbor, MI, 48109}
    \address[Johns Hopkins Bloomberg School of Public Health]{Department of Biostatistics, 615 North Wolfe Street, Baltimore, MD,
21205}

@@
